---
title: "Lab 1"
author: "Taylor Wise"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

Tasks for lab 1

## Task 1

Obtain the working directory
```{r}
getwd()
```


## Task 2
Read the data and obtain the first 6 lines of the data
```{r}
ddt <- read.csv("DDT.csv")
head(ddt)
```

## Task 3

```{r}
names(ddt)
```

RIVER, SPECIES = Qualitative

MILE, LENGTH, WEIGHT, DDT = Quantitative

```{r}
tab <- table(ddt$SPECIES)
tab
```

Number of Species = 3

The species are CCATFISH, LMBASS, and SMBUFFALO

Subset the ddt data set so that we see only the observations with SPECIES, LMBASS and WEIGHT bigger than 800gms
```{r}
ddt[ddt$SPECIES == "LMBASS" & ddt$WEIGHT > 800,]
```


Subset the ddt data set so that we see only those observations where the RIVER is SCM and DDT bigger than 4.0
```{r}
ddt[ddt$RIVER == "SCM" & ddt$DDT > 4.0,]
```

## Task 4

Mean length of fish
```{r}
mean(ddt$LENGTH)
```

Mean length of fish is 42.8125
<!--
L <- ddt$LENGTH
mean(L)

Same result
-->

Standard deviation of the weight of fish
```{r}
sd(ddt$WEIGHT)
```

Standard deviation of the weight of fish is 376.5461
<!--
with(ddt,plot(WEIGHT,LENGTH))

is the same thing as below basically (no color ofc)
-->

Graph of LENGTH vs WEIGHT
```{r}
plot(ddt$WEIGHT, ddt$LENGTH, xlab="WEIGHT", ylab="LENGTH", col="red")
```

0 

No, the graph provided in the lab document is not the graph of LENGTH vs WEIGHT because it looks different than the actual graph provided above

<!--
Y versus X s.t. Length is the y-axis and WEIGHT is the x-axis  
-->

If v=1:20, what is the last value of v/20?
```{r}
v=1:20
print(v/20)
```

## Task 5

Table of the rivers
```{r}
table(ddt$RIVER)
```

<!--
Aesthetic colors for the graphs
-->

Color coded barplot of the rivers
```{r}
barplot(table(ddt$RIVER), col = c("darkseagreen3", "lightsteelblue3", "burlywood3", "rosybrown3"))
```

Table of the rivers crossed with the species of fish
```{r}
table(ddt$SPECIES,ddt$RIVER)
```

Barplot of the rivers crossed with the species of fish
```{r}
barplot(table(ddt$SPECIES,ddt$RIVER), col = c("lightsteelblue3", "burlywood3", "rosybrown3"), beside=TRUE)
```

## Task 6

Pie chart of fish SPECIES
```{r}
pie(table(ddt$SPECIES), col = c("rosybrown3", "burlywood3", "lightsteelblue3"))
```

Pie chart of RIVERS
```{r}
pie(table(ddt$RIVER), col = c("darkseagreen3", "rosybrown3", "burlywood3", "lightsteelblue3"))
```

## Task 7

Boxplot of DDT
```{r}
boxplot(ddt$DDT, data=ddt)
```

Boxplot of WEIGHT
```{r}
boxplot(ddt$WEIGHT, data=ddt)
```

Boxplot of LENGTH
```{r}
boxplot(ddt$LENGTH, data=ddt)
```

## Task 8

Coplot of LENGTH versus WEIGHT given RIVER
```{r}
coplot(LENGTH~WEIGHT|RIVER, data=ddt, col=c("darkseagreen3", "rosybrown3", "burlywood3", "lightsteelblue3"))
```

Coplot of DDT versus WEIGHT given SPECIES
```{r}
coplot(DDT~WEIGHT|SPECIES, data=ddt, col=c("darkseagreen3", "rosybrown3", "burlywood3", "lightsteelblue3"))
```

## Task 9

Boxplot
```{r}
library(ggplot2)
ggplot(ddt, aes(x=SPECIES, y=WEIGHT, fill=RIVER,)) + geom_boxplot(col="black") + ggtitle("Taylor Wise")
```

Violin Plot
```{r}
library(ggplot2)
ggplot(ddt, aes(x=RIVER, y=LENGTH, fill=SPECIES,)) + geom_violin(col="black") + ggtitle("Taylor Wise")
```

Scatter plot
```{r}
library(ggplot2)
ggplot(ddt, aes(x=WEIGHT, y=LENGTH, color=SPECIES)) + geom_point() + ggtitle("Taylor Wise")

```



