---
title: "Lab2"
author: "Taylor Wise"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1
Get the working directory
```{r}
getwd()
```

## Task 2
Read in data and obtain first 6 lines of data
```{r}
epagas <- read.csv("EPAGAS.csv")
head(epagas)
```

## Task 3
Make the object mpg, the number of miles per gallon vector
```{r}
mpg = epagas$MPG
head(mpg)
```

Transform the mpg variable to z
```{r}
z = (mpg - mean(mpg)) / sd(mpg)
head(z)
```

Verify z
```{r}

```

Possible outliers
```{r}
mpg[abs(z) >= 2 & abs(z) <= 3]
```

Actual outliers
```{r}
mpg[abs(z) > 3]
```

Dotplot
```{r}
mpgcol = ifelse(abs(z) > 3, "red", ifelse(abs(z) >= 2 & abs(z) <= 3, "blue", "black"))

library(lattice)
dotplot(mpg, col = mpgcol, main = "Plot of mpg values")
```

## Task 4
Boxplot of mpg
```{r}
boxplot(mpg, main = "MPG Boxplot", xlab = "MPG", col = "seagreen3", horizontal = TRUE, notch = TRUE)
```

Chebyshev's Theorem claims that at least 75% of the data are between 2 standard deviations
```{r}
length(mpg[abs(z) < 2]) / length(mpg)
```

As seen above with a proportion of 0.96, at least 75% of the data are between 2 standard deviations which aggrees with Chebyshev

Empirical Rule
```{r}

```

